### 2020-03-05 Common/chapters/literature_review.tex
Julien and Johan: patient split figure + global performance + minors

\setlength{\marginparwidth}{3cm}\leavevmode \marginnote{\textbf{Johan}}Since this work focuses on lesions classification, this chapter presents the main research works about CADx systems based on convolutional neural networks for each body part used in sections \ref{ch:paper_reproduction} and \ref{ch:transfer_learning}, i.e. prostate, lung and brain. Furthermore, it exclusively focuses on studies that used the same datasets as the ones used in this work.

>>>>

\setlength{\marginparwidth}{3cm}\leavevmode \marginnote{\textbf{Johan}}Since this work focuses on lesion classification, this chapter presents the main research works about CAD systems based on convolutional neural networks for each body part used in sections \ref{ch:paper_reproduction} and \ref{ch:transfer_learning}, i.e. prostate, lung and brain. Furthermore, it exclusively focuses on studies that used the same datasets as the ones used in this work.

### 2020-03-05 Common/chapters/paper_reproduction.tex
Julien and Johan: patient split figure + global performance + minors

\subsubsection{Methodology}
\setlength{\marginparwidth}{3cm}\leavevmode \marginnote{\textbf{Julien}}
The methodology described in this section is the closest replication possible of the authors' data processing methodology. Most of it were reproductible in a similar way, apart from the manual lesion contouring which was performed by a qualified radiologist in the authors' case. Furthermore, the authors did not mention any participation in the official PROSTATEx challenge. In order to get an unbiased evaluation of the performance of our model, these images were also processed as explained below in order to take part in the challenge, but were not augmented. Nevertheless, the authors created their own test set by splitting the PROSTATEx training images into a training set ($80\%$), a validation set ($10\%$) and a test set ($10\%$). They mentioned that the test set was augmented $11$ times but gave no information about the training and validation sets. Since the training set is imbalanced (3 false for 1 true), the true class was augmented more times than the false class ($60$x) in order to create balanced training and validation sets.

@@

Finally, the data was split into three subsets (training ($80$\%), validation ($10$\%) and test ($10$\%)) and was exported as NumPy arrays. The percentages represent a percentage of patients. Figure \ref{fig:paper_reproduction_split} explains the entire splitting process and gives details about the patients that are in each set. In fact, each subset contains respectively 80\%, 10\% and 10\% of all patients belonging to the "true" and "false" class. At the beginning of the process, class "false" contains $\lfloor 0.8 * 164 \rfloor = 131$ training patients, $\lfloor 0.1 * 164 \rfloor = 16$ test patients and $164 - 131 - 16 = 17$ validation patients. Class "true" contains $\lfloor 0.8 * 69 \rfloor = 55$ training patients, $\lfloor 0.1 * 69 \rfloor = 6$ test patients and $69 - 55 - 6 = 8$ validation patients. Then, during the stacking process, some patients are not used since the three sequences are not available for them. These patients are simply discarded. At the end, the "false" class is composed of 131 training patients, 17 validation patients and 15 test patients, whereas the "true" class is composed of 55 training patients, 8 validation patients and 4 test patients. This approach does not take the number of images into account (a patient may have multiple lesions for example) but guarantees that every lesion is exclusive to a specific subset. In other words, it is impossible to find a slighty different version of the same lesion (resulting from the data augmentation) in the training set and in the validation set for example. This would be problematic since the validation set, as well as the test set, need to contain unseen data to be able to evaluate the generalization ability of the model in an unbiased way.

\begin{figure}[!h]
\centering
\includegraphics[width=1\textwidth, keepaspectratio=true]{./figures/paper_reproduction_split.png}
\caption{Splitting process}
\label{fig:paper_reproduction_split}
\end{figure}

>>>>

\subsubsection{Methodology}
\setlength{\marginparwidth}{3cm}\leavevmode \marginnote{\textbf{Julien}}The methodology described in this section is the closest replication possible of the authors' data processing methodology. Most of it were reproductible in a similar way, apart from the manual lesion contouring which was performed by a qualified radiologist in the authors' case. Furthermore, the authors did not mention any participation in the official PROSTATEx challenge. In order to get an unbiased evaluation of the performance of our model, these images were also processed as explained below in order to take part in the challenge, but were not augmented. Nevertheless, the authors created their own test set by splitting the PROSTATEx training images into a training set ($80\%$), a validation set ($10\%$) and a test set ($10\%$). They mentioned that the test set was augmented $11$ times but gave no information about the training and validation sets. Since the training set is imbalanced (3 false for 1 true), the true class was augmented more times than the false class (which was augmented $60$x) in order to create balanced training and validation sets.

@@

Finally, the data was split into three subsets (training ($80$\%), validation ($10$\%) and test ($10$\%)) and was exported as NumPy arrays. The percentages represent a percentage of patients. Figure \ref{fig:paper_reproduction_split} explains the entire splitting process and gives details about the patients that are in each set. In fact, each subset contains respectively 80\%, 10\% and 10\% of all patients belonging to the "true" and "false" class. At the beginning of the process, class "false" contains $\lfloor 0.8 * 164 \rfloor = 131$ training patients, $\lfloor 0.1 * 164 \rfloor = 16$ test patients and $164 - 131 - 16 = 17$ validation patients. On the other hand, class "true" contains $\lfloor 0.8 * 69 \rfloor = 55$ training patients, $\lfloor 0.1 * 69 \rfloor = 6$ test patients and $69 - 55 - 6 = 8$ validation patients. Then, during the stacking process, some patients were not used since the three sequences were not available for them. These patients were simply discarded. At the end, the "false" class was composed of 131 training patients, 17 validation patients and 15 test patients, whereas the "true" class was composed of 55 training patients, 8 validation patients and 4 test patients. This approach does not take the number of images into account (a patient may have multiple lesions for example) but guarantees that every lesion is exclusive to a specific subset. In other words, it is impossible to find a slighty different version of the same lesion (resulting from the data augmentation) in the training set and in the validation set for example. This would be problematic since the validation set, as well as the test set, need to contain unseen data to be able to evaluate the generalization ability of the model in an unbiased way.

\begin{figure}[!h]
\centering
\includegraphics[width=1\textwidth, keepaspectratio=true]{./figures/paper_reproduction_split.png}
\caption{Patient splitting process}
\label{fig:paper_reproduction_split}
\end{figure}

### 2020-03-05 Common/chapters/transfer_learning.tex
Julien and Johan: patient split figure + global performance + minors

\subsection{Visualization of the impact of the various datasets on the target task}
\label{sec:global_performance}
\setlength{\marginparwidth}{3cm}\leavevmode \marginnote{\textbf{Julien}}On paper, each step of the transfer learning is supposed to make the model learn new features that are useful for the first dataset (the target dataset). As described in section \ref{sec:process_overview}, the model can be seen as a feature extractor and a decision maker. Hence, to see how much each step was actually helping the process, the decision maker resulting from the first step of the transfer learning, i.e. the training on the first/target dataset, was saved. As this decision maker is saved after the training on the target dataset, it is specialized in the classification of the latter. In other words, if the target dataset is a prostate dataset, it will classify prostate tumors with a relatively high precision.

Then, after each epoch of every other transfer learning step, the saved decision maker was attached to the current feature extractor. This temporary model was then tested on the target training and validation sets, which resulted in different training and validation metrics.

At the end of the transfer learning pipeline, a graph was generated for each metric. These curves show the evolution of the latter across the entire process, which allows to determine the impact of each non-target dataset on the target task. If a metric is increasing, the features learned thanks to the corresponding dataset are useful for the target one. Figure \ref{fig:global_performance_example} shows an example of this graph for the AUC.\\
As expected, the curve is flat during the frozen parts. Since the same decision maker is attached to frozen layers (which remain the same by definition), the values of the metrics do not change at all.

\begin{figure}[!h]
\centering
\includegraphics[width=1\textwidth, keepaspectratio=true]{./figures/global_performance_example.png}
\caption{Please replace this figure with a real example as soon as the results are available!}
\label{fig:global_performance_example}
\end{figure}

@@

\subsection{Global performance}
\setlength{\marginparwidth}{3cm}\leavevmode \marginnote{\textbf{Julien}}The so-called global performance graphs are a visual representation of the impact of each part of the transfer learning and each dataset on the target task (see Section \ref{sec:global_performance}). In this case, they show how much the features learned from each dataset help classify prostate lesions. As a reminder, a temporary model was created and tested at the end of each epoch. The latter was composed of the current feature extractor attached to the decision maker of the best model resulting from the first training on the prostate dataset.

Figure \ref{fig:tl_global_validation_auc} shows the evolution of the AUC on the validation set of PROSTATEx. As expected, the first training on PROSTATEx generated the same curve as on \ref{fig:tl_ds1_full}). The latter looks steeper because of the difference in the X-axis scale.\\
Then, the frozen training on the second dataset (as well as the other frozen trainings) shows a flat line. Reason for that is that the feature extractor and the decision maker do not change here. In fact, the former is frozen and the latter is always the same. As the decision maker is the one coming from the best model of the first training according to the validation AUC, the flat line is as high as the best validation AUC value reached during the first training. \\
Furthermore, the full training curve shows that brain cancer features are actually helpful. If they were not, the metric would not have increased as fast and would not have reached such values.\\
During the full training on the lung dataset, the AUC starts at a high value after the first epoch. The combination of the prostate, brain and lung features seem to help the model generalize. This hypothesis is confirmed by the values reached during last training on the prostate dataset. In fact, the AUC value is higher than the one reached during the first training. Interestingly, the curve starts high, improves for 50 epochs and starts decreasing. This implies that the model reached a high level of generalization during transfer learning. These 50 epochs were enough to keep the low level features learned thanks to the other datasets while adapting the model to the target task again.

@@

\begin{figure}[!h]
\centering
\noindent
\makebox[\textwidth]{
\includegraphics[width=1\textwidth, keepaspectratio=true]{./figures/tl_global_validation_auc.png}
}
\caption{Transfer learning - Global Performance - Validation AUC}
\label{fig:tl_global_validation_auc.png}
\end{figure}

>>>>

%\subsection{Visualization of the impact of the various datasets on the target task}
%\label{sec:global_performance}
%\setlength{\marginparwidth}{3cm}\leavevmode \marginnote{\textbf{Julien}}On paper, each step of the transfer learning is supposed to make the model learn new features that are useful for the first dataset (the target dataset). As described in section \ref{sec:process_overview}, the model can be seen as a feature extractor and a decision maker. Hence, to see how much each step was actually helping the process, the decision maker resulting from the first step of the transfer learning, i.e. the training on the first/target dataset, was saved. As this decision maker is saved after the training on the target dataset, it is specialized in the classification of the latter. In other words, if the target dataset is a prostate dataset, it will classify prostate tumors with a relatively high precision. 
%
%Then, after each epoch of every other transfer learning step, the saved decision maker was attached to the current feature extractor. This temporary model was then tested on the target training and validation sets, which resulted in different training and validation metrics. 
%
%At the end of the transfer learning pipeline, a graph was generated for each metric. These curves show the evolution of the latter across the entire process, which allows to determine the impact of each non-target dataset on the target task. If a metric is increasing, the features learned thanks to the corresponding dataset are useful for the target one. Figure \ref{fig:global_performance_example} shows an example of this graph for the AUC.\\
%As expected, the curve is flat during the frozen parts. Since the same decision maker is attached to frozen layers (which remain the same by definition), the values of the metrics do not change at all. 
%
%\begin{figure}[!h]
%\centering
%\includegraphics[width=1\textwidth, keepaspectratio=true]{./figures/global_performance_example.png}
%\caption{Please replace this figure with a real example as soon as the results are available!}
%\label{fig:global_performance_example}
%\end{figure}

@@

\subsection{Visualization of the impact of the various datasets on the target task}
\setlength{\marginparwidth}{3cm}\leavevmode \marginnote{\textbf{Julien}}The so-called global performance graphs are a visual representation of the impact of each part of the transfer learning and each dataset on the target task. In this case, they show how much the features learned from each dataset help classify prostate lesions. On paper, each step of the transfer learning is supposed to make the model learn new features that are useful for the first dataset (the target dataset). As described in section \ref{sec:process_overview}, the model can be seen as a feature extractor and a decision maker. Hence, to see how much each step was actually helping the process, the decision maker resulting from the best model of the first step of the transfer learning, i.e. the training on the first/target dataset, was saved. As this decision maker is saved after the training on the target dataset, it is specialized in the classification of the latter. In other words, if the target dataset is a prostate dataset, it will classify prostate tumors with a relatively high precision.\\
Then, after each epoch of every other transfer learning step, the saved decision maker was attached to the current feature extractor. This temporary model was then tested on the target training and validation sets, which resulted in different training and validation metrics.\\
At the end of the transfer learning pipeline, a graph was generated for each metric. These curves show the evolution of the latter across the entire process, which allows to determine the impact of each non-target dataset on the target task. If a metric is increasing, the features learned thanks to the corresponding dataset are useful for the target one.

Figure \ref{fig:tl_global_validation_auc} shows the evolution of the AUC on the validation set of PROSTATEx. As expected, the first training on PROSTATEx generated the same curve as on Figure \ref{fig:tl_ds1_full}. The latter looks steeper because of the difference in the X-axis scale.\\
Moreover, the frozen training on the second dataset (as well as the other frozen trainings) shows a flat line. Reason for that is that the feature extractor and the decision maker do not change here. In fact, the former is frozen and the latter is always the same. As the decision maker is the one coming from the best model of the first training according to the validation AUC, the flat line is as high as the best validation AUC value reached during the first training.

Furthermore, the full training curve shows that brain cancer features are actually helpful. If they were not, the metric would not have increased as fast and would not have reached such values.

During the full training on the lung dataset, the AUC starts at a high value after the first epoch. The combination of the prostate, brain and lung features seem to help the model generalize. This hypothesis is confirmed by the values reached during last training on the prostate dataset. In fact, the AUC value is higher than the one reached during the first training. Interestingly, the curve starts high, improves for 50 epochs and starts decreasing. This implies that the model reached a high level of generalization during transfer learning. These 50 epochs were enough to keep the low level features learned thanks to the other datasets while adapting the model to the target task again.

%\subsection{Visualization of the impact of the various datasets on the target task}
%\label{sec:global_performance}
%\setlength{\marginparwidth}{3cm}\leavevmode \marginnote{\textbf{Julien}}On paper, each step of the transfer learning is supposed to make the model learn new features that are useful for the first dataset (the target dataset). As described in section \ref{sec:process_overview}, the model can be seen as a feature extractor and a decision maker. Hence, to see how much each step was actually helping the process, the decision maker resulting from the first step of the transfer learning, i.e. the training on the first/target dataset, was saved. As this decision maker is saved after the training on the target dataset, it is specialized in the classification of the latter. In other words, if the target dataset is a prostate dataset, it will classify prostate tumors with a relatively high precision. 
%
%Then, after each epoch of every other transfer learning step, the saved decision maker was attached to the current feature extractor. This temporary model was then tested on the target training and validation sets, which resulted in different training and validation metrics. 
%
%At the end of the transfer learning pipeline, a graph was generated for each metric. These curves show the evolution of the latter across the entire process, which allows to determine the impact of each non-target dataset on the target task. If a metric is increasing, the features learned thanks to the corresponding dataset are useful for the target one. Figure \ref{fig:global_performance_example} shows an example of this graph for the AUC.\\
%As expected, the curve is flat during the frozen parts. Since the same decision maker is attached to frozen layers (which remain the same by definition), the values of the metrics do not change at all.

@@

\begin{figure}[!h]
\centering
\noindent
\makebox[\textwidth]{
\includegraphics[width=1\textwidth, keepaspectratio=true]{./figures/tl_global_validation_auc.png}
}
\caption{Transfer learning - Global Performance - Validation AUC}
\label{fig:tl_global_validation_auc}
\end{figure}
